On exponential sums arising from the classical doubling method

### 2025-03-29 main.tex

	The following theorem summarizes these cases.
	\begin{theorem}
	A maximal rational torus $\algebraicGroup{T}$ of $\algebraicGroup{G} =  \algebraicGroup{\GSp}_{2n}, \algebraicGroup{\GSO}_{2n}^{\pm}$ is isomorphic to \begin{align*}
				^{L^{-1}w}\algebraicGroup{T}_0 \cong& \left\{\left(x,y,z\right) \in  \restrictionOfScalars{\finiteFieldExtension{\lambda^+}}{\finiteField}{\multiplcativeScheme} \times \restrictionOfScalars{\finiteFieldExtension{2\lambda^-}}{\finiteField}{\multiplcativeScheme}  \times \multiplcativeScheme \mid \FieldNorm{2\lambda^-}{\lambda^-}(y) = z \right\}.
	\end{align*}
	where $\left(\lambda^+, \lambda^-\right)$ is a pair of partitions such that $\sizeof{\lambda^+} + \sizeof{\lambda^-} = n$. We have that $$\algebraicGroup{T}^{\Frobenius} \cong \left\{\left(x, y, z\right) \in \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \times \multiplicativegroup{\finiteFieldExtension{2 \lambda^-}} \times \multiplicativegroup{\finiteField} \mid \FieldNorm{2 \lambda^{-}}{\lambda^{-}}\left(y\right) = z \right\}.$$ The embedding $\algebraicGroup{T}^{\Frobenius} \to \algebraicGroup{G}^{\Frobenius}$ is of the form $$\left(x, y, z\right) \mapsto
	\begin{pmatrix}
		z \cdot \iota_+ \left(x\right)\\
		& \iota_{-}\left(y\right) &\\
		& & \iota_+^{\ast}\left(x\right)
	\end{pmatrix},$$
	where $\iota_+ \colon \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \to \GL_{\sizeof{\lambda^+}}\left(\finiteField\right)$ and $$\iota_- \colon \left\{ \left(y, z\right)\in \multiplicativegroup{\finiteFieldExtension{2 \lambda^-}} \times \multiplicativegroup{\finiteField} \mid \FieldNorm{2 \lambda^{-}}{\lambda^{-}}\left(y\right) = z \text{ for all } i \right\} \to G'$$ are embeddings. Here, $G'$ is a group of the same type as $\algebraicGroup{G}^{\Frobenius}$, with $n$ replaced by $\sizeof{\lambda^{-}}$.
\end{theorem}

@@

\subsection{Multiplicativity property}
The goal of this section is to understand how the doubling method Jacobi sums behave under parabolic induction.

@@

\subsection{Computation of character sums attached to Deligne--Lusztig data}

@@

\subsection{Invariance under geometric conjugacy}
\calvin{todo:type in some motivation for class functions whose product doesn't depend on geometric conjugacy, gamma functions whatever}.

@@

\subsection{Computation for geometric Lusztig series}

@@

\calvin{end alternate writeup}
\end{comment}
\calvin{start alternate writeup 2 of section 3.4/3.4.1}
\subsection{Computation for Lusztig Series}
The goal of this section is to prove \Cref{thm:Phi_stable} in particular:
\begin{theorem}
\label{thm:doubling-method-gamma-factor-for-deligne-lusztig} For either $G=G$ or its extension $\GroupExtension{G}$. If $c_{\pi}$ is nonzero in the virtual representation $R_T(\theta) = \sum c_{\pi} \pi$ then
$$\dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = g_{T \cap G}\left(\theta|_{T \cap G}, \fieldCharacter\right)$$
\end{theorem}
\calvin{add in $\chi$} the right hand side of which we showed in \Cref{lem:indep_geo_conj} only depends on the geometric conjugacy class of $(T,\theta)$ hence on the Lusztig series $\LusztigSeries{G}{s}$ which contains $\pi$, implying \Cref{thm:Phi_stable}.
\subsubsection{Proof of \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig}}
We first reduce \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} to the case $G$ has connected center. For any irreducible representation $\pi$ of $G$ which is the a component of an irreducible representation $\tilde{\pi}$ of $\GroupExtension{G}$ restricted to $\pi$,
\[
    \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \dblGammaFactorSpace{\hermitianSpace}{\tilde{\pi}}{\chi}{\fieldCharacter}.
\]

>>>>

	The following theorem summarizes these cases.
	\begin{theorem}
	A maximal rational torus $\algebraicGroup{T}$ of $\algebraicGroup{G} =  \algebraicGroup{\GSp}_{2n}, \algebraicGroup{\GSO}_{2n}^{\pm}$ is isomorphic to \begin{align*}
				^{L^{-1}w}\algebraicGroup{T}_0 \cong& \left\{\left(x,y,z\right) \in  \restrictionOfScalars{\finiteFieldExtension{\lambda^+}}{\finiteField}{\multiplcativeScheme} \times \restrictionOfScalars{\finiteFieldExtension{2\lambda^-}}{\finiteField}{\multiplcativeScheme}  \times \multiplcativeScheme \mid \FieldNorm{2\lambda^-}{\lambda^-}(y) = z \right\}.
	\end{align*}
	where $\left(\lambda^+, \lambda^-\right)$ is a pair of partitions such that $\sizeof{\lambda^+} + \sizeof{\lambda^-} = n$. We have that $$\algebraicGroup{T}^{\Frobenius} \cong \left\{\left(x, y, z\right) \in \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \times \multiplicativegroup{\finiteFieldExtension{2 \lambda^-}} \times \multiplicativegroup{\finiteField} \mid \FieldNorm{2 \lambda^{-}}{\lambda^{-}}\left(y\right) = z \right\}.$$ The embedding $\algebraicGroup{T}^{\Frobenius} \to \algebraicGroup{G}^{\Frobenius}$ is of the form $$\left(x, y, z\right) \mapsto
	\begin{pmatrix}
		z \cdot \iota_+ \left(x\right)\\
		& \iota_{-}\left(y,z\right) &\\
		& & \iota_+^{\ast}\left(x\right)
	\end{pmatrix},$$
	where $\iota_+ \colon \multiplicativegroup{\finiteFieldExtension{\lambda^+}} \to \GL_{\sizeof{\lambda^+}}\left(\finiteField\right)$ and $$\iota_- \colon \left\{ \left(y, z\right)\in \multiplicativegroup{\finiteFieldExtension{2 \lambda^-}} \times \multiplicativegroup{\finiteField} \mid \FieldNorm{2 \lambda^{-}}{\lambda^{-}}\left(y\right) = z \text{ for all } i \right\} \to G'$$ are embeddings. Here, $G'$ is a group of the same type as $\algebraicGroup{G}^{\Frobenius}$, with $n$ replaced by $\sizeof{\lambda^{-}}$.
\end{theorem}

@@

\subsection{Multiplicativity}
The goal of this section is to understand how the doubling method Jacobi sums behave under parabolic induction.

@@

\subsection{Pairing with Deligne--Lusztig characters}
\calvin{old title: Computation of character sums attached to Deligne--Lusztig data}

@@

\subsection{Invariance of $\langle \JacobiKernel{}, R_T(\theta) \rangle$ under geometric conjugacy}

@@

\subsection{Computation for geometric Lusztig series}
\calvin{retitle: Stability of Jacobi sums}

@@

\calvin{end alternate writeup}
\end{comment}
\calvin{start alternate writeup 2 of section 3.4/3.4.1}
\subsection{Stability}
The goal of this section is to prove \Cref{thm:Phi_stable} in particular:
\begin{theorem}
\label{thm:doubling-method-gamma-factor-for-deligne-lusztig} For either $G=G$ or its extension $\GroupExtension{G}$. If $c_{\pi}$ is nonzero in the virtual representation $R_T(\theta) = \sum c_{\pi} \pi$ then
$$\dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = g_{T \cap G}\left(\theta|_{T \cap G}, \fieldCharacter\right)$$
\end{theorem}
\calvin{add in $\chi$} the right hand side of which we showed in \Cref{lem:indep_geo_conj} only depends on the geometric conjugacy class of $(T,\theta)$ hence on the Lusztig series $\LusztigSeries{G}{s}$ which contains $\pi$, implying \Cref{thm:Phi_stable}.
\subsubsection{Proof of \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig}}
We first reduce \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} to the case $G$ has connected center. For any irreducible representation $\pi$ of $G$ which is the a component of an irreducible representation $\tilde{\pi}$ of $\GroupExtension{G}$ restricted to $\pi$,
\[
    \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \dblGammaFactorSpace{\hermitianSpace}{\tilde{\pi}}{\chi}{\fieldCharacter}.
\]
